\bf A Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Electrons

### 2025-12-10 submissions/Earlier Versions of Manuscript/tilingProject.tex
started creating referee response

The subsystem \(S\) is usually taken to be the "cluster", and consequently, \(R\) represents the "bath".
The smaller system is typically chosen such that its eigenstates are known exactly.
Progress is then made by choosing a simpler version of \(H_R\) and a simpler form also for its coupling \(H_{RS}\) with the smaller system.
This combination of the cluster and the simpler bath is then called the \textit{auxiliary system}.
A typical auxiliary system for the Hubbard model could, potentially, be the single-impurity Anderson model (SIAM), where the impurity represents an arbitrary site \(p\) of the lattice, the bath represents the rest of the lattice sites and the hybridisation term between the impurity and the bath represents the coupling term \(H_{RS}\):
\begin{equation}\begin{aligned}\label{clus_bath_siam}
	\mathcal{H}_\text{SIAM} = \underbrace{\epsilon_d \hat n_d + U \hat n_{d \uparrow} \hat n_{d \downarrow}}_{H_S} + \underbrace{- t\sum_{\left<i,j \right>, \sigma}\left(c^\dagger_{i\sigma}c_{j\sigma} + \text{h.c.}\right)}_{H_{R}} + \underbrace{V\sum_\sigma \left( c^\dagger_{d\sigma}c_{0\sigma} + \text{h.c.}\right)}_{H_{SR}}~.
\end{aligned}\end{equation}
Such a construction is shown in fig.~\ref{cluster-bath}.
\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.48\textwidth]{clusterBath.pdf}
	\caption{\textit{Left}: Full Hubbard model lattice with onsite repulsion $U^H$ on all sites and hopping between nearest neighbour sites with strength $t^H$. \textit{Right}: Extraction of the auxiliary (cluster+bath) system from the full lattice. The central site on left becomes the impurity site (red) on the right (with an onsite repulsion $\epsilon_d$), while the rest of the $N-1$ sites on the left form a conduction bath (green circles) (with dispersion $\epsilon_k$ and correlation modelled by the self-energy $\Sigma_k(\omega)$) that hybridizes with the impurity through the coupling $V$.}
	\label{cluster-bath}
\end{figure}
\textit{It should be noted that any reasonable choice of the cluster and bath would break the translational symmetry of the full model}. To allow computing quantities, one would need to make the bath (which is a much larger system) simpler than the cluster (which is a single site). This distinction breaks the translational symmetry of the Hubbard model. For eg., if one chooses eq.~\ref{clus_bath_siam} as the auxiliary system, then the fact that the impurity has an onsite correlation while the bath does not means we have broken the symmetry between the cluster and the bath.

@@

In the present work, the goal is to tackle interacting fermions on a 2D square lattice, which makes it important for us to incorporate the effects of the lattice within the impurity model. More specifically, while the impurity-bath interactions on the infinite dimensional Bethe lattice are \(s-\)wave symmetric, that is not a realistic form for a 2D lattice. To make a realistic impurity model, we embed the impurity site into the lattice of a 2D conduction bath, and lower the \(s-\)wave symmetry of the interactions down to the \(C_4\) symmetry of the 2D square lattice. A general structure of such a model is shown in fig.~\ref{embeddedEsiam}; one of the sites of the 2D square lattice is identified as the impurity (red), while the four nearest-neghbours (navy blue) interact directly with this impurity site through a \(C_4\)-symmetric interaction.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.2\textwidth]{pWaveEsiam.pdf}
	\caption{General structure of an embedded extended SIAM, where the impurity site is a part of the 2D square lattice. The four nearest-neghbour sites (dark blue) interact with the impurity site through one or more processes, and can host local interactions. The other lattice sites (light blue) are completely non-interacting.}
	\label{embeddedEsiam}
\end{figure}
\subsection{Hamiltonian}
We consider an impurity spin \(\vec S_d\) interacting with a two-dimensional tight-binding conduction bath through a highly localised interaction, described by the Hamiltonian
\begin{equation}\begin{aligned}
	\mathcal{H} = H_\text{cbath} + H_\text{imp-cbath} + H_\text{cbath-int}~,
\end{aligned}\end{equation}
where \(H_\text{cbath}\) is the kinetic energy arising out of nearest-neighbour hopping processes,
\begin{equation}\begin{aligned}
	H_\text{cbath} = -2t\sum_{{\bf k}}\left[\cos(ak_x) + \cos(ak_y)\right] c^\dagger_{{\bf k},\sigma}c_{{\bf k},\sigma}~.
\end{aligned}\end{equation}
For the impurity-bath interaction \(H_\text{imp-cbath}\) and the local interaction \(H_\text{cbath-int}\) in the bath, we take the simplest choices consistent with the 2D square lattice geometry:
\begin{equation}\begin{aligned}
	H_\text{imp-cbath} &=\frac{1}{2} J\sum_{\sigma_1,\sigma_2}\sum_{Z} {\bf S}_d\cdot c^\dagger_{Z\sigma_1}{\boldsymbol \tau}_{\sigma_1,\sigma_2} c_{Z\sigma_2}~,\\
	H_\text{cbath-int} &= -\frac{W}{2}\sum_{Z} \left(n_{Z\uparrow} - n_{Z\downarrow}\right)^2 ~,
\end{aligned}\end{equation}
where \(Z\) sums over all nearest-neighbour sites of the impurity site (navy blue circles in Fig.~\ref{embeddedEsiam}, henceforth referred to as bath zeroth sites). The impurity interacts with the bath zeroth sites in \(H_\text{imp-cbath}\) through four local Kondo terms, of uniform strength. The other term \(H_\text{cbath-int}\) represents local Hubbard interactions on the four zeroth sites, again of uniform strength.

The 2D structure of the impurity model and its distinction from the infinite-dimensional counterpart becomes apparent once we transform the Hamiltonian to momentum space:
\begin{equation}\begin{aligned}
	J_{{\bf k}, {\bf k}^\prime} &= \frac{J}{2}\left[\cos\left({\bf k}_x - {\bf k}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y\right)\right] ~,\\
	W_{{\bf k}, {\bf k}^\prime,{\bf q}, {\bf q}^\prime} &= W\left[\cos\left({\bf k}_x - {\bf k}^\prime_x + {\bf q}_x - {\bf q}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y + {\bf q}_y - {\bf q}^\prime_y\right)\right] ~.
\end{aligned}\end{equation}
where the subscripts \(x\) and \(y\) represent components of the momenta vector. This form shows that the Kondo coupling is no longer independent of the exchange momentum, and encodes information of the lattice geometry in it.

>>>>

The subsystem \(S\) is usually taken to be the "cluster", and consequently, \(R\) represents the "bath".
The smaller system is typically chosen such that its eigenstates are known exactly.
Progress is then made by choosing a simpler version of \(H_R\) and a simpler form also for its coupling \(H_{RS}\) with the smaller system.
This combination of the cluster and the simpler bath is then called the \textit{auxiliary system}.
A typical auxiliary system for the Hubbard model could, potentially, be the single-impurity Anderson model (SIAM), where the impurity represents an arbitrary site \(p\) of the lattice, the bath represents the rest of the lattice sites and the hybridisation term between the impurity and the bath represents the coupling term \(H_{RS}\):
\begin{equation}\begin{aligned}\label{clus_bath_siam}
	\mathcal{H}_\text{SIAM} = \underbrace{\epsilon_d \hat n_d + U \hat n_{d \uparrow} \hat n_{d \downarrow}}_{H_S} + \underbrace{- t\sum_{\left<i,j \right>, \sigma}\left(c^\dagger_{i\sigma}c_{j\sigma} + \text{h.c.}\right)}_{H_{R}} + \underbrace{V\sum_\sigma \left( c^\dagger_{d\sigma}c_{0\sigma} + \text{h.c.}\right)}_{H_{SR}}~.
\end{aligned}\end{equation}
Such a construction is shown in fig.~\ref{cluster-bath}.
\begin{figure}[!htb]
	\centering

    \includegraphics[width=0.48\textwidth]{clusterBath.pdf}
	\caption{\textit{Left}: Full Hubbard model lattice with onsite repulsion $U^H$ on all sites and hopping between nearest neighbour sites with strength $t^H$. \textit{Right}: Extraction of the auxiliary (cluster+bath) system from the full lattice. The central site on left becomes the impurity site (red) on the right (with an onsite repulsion $\epsilon_d$), while the rest of the $N-1$ sites on the left form a conduction bath (green circles) (with dispersion $\epsilon_k$ and correlation modelled by the self-energy $\Sigma_k(\omega)$) that hybridizes with the impurity through the coupling $V$.}
	\label{cluster-bath}
\end{figure}
\textit{It should be noted that any reasonable choice of the cluster and bath would break the translational symmetry of the full model}. To allow computing quantities, one would need to make the bath (which is a much larger system) simpler than the cluster (which is a single site). This distinction breaks the translational symmetry of the Hubbard model. For eg., if one chooses eq.~\ref{clus_bath_siam} as the auxiliary system, then the fact that the impurity has an onsite correlation while the bath does not means we have broken the symmetry between the cluster and the bath.

@@

In the present work, the goal is to tackle interacting fermions on a 2D square lattice, which makes it important for us to incorporate the effects of the lattice within the impurity model. More specifically, while the impurity-bath interactions on the infinite dimensional Bethe lattice are \(s-\)wave symmetric, that is not a realistic form for a 2D lattice. To make a realistic impurity model, we embed the impurity site into the lattice of a 2D conduction bath, and lower the \(s-\)wave symmetry of the interactions down to the \(C_4\) symmetry of the 2D square lattice. A general structure of such a model is shown in fig.~\ref{embeddedEsiam}; one of the sites of the 2D square lattice is identified as the impurity (red), while the four nearest-neghbours (navy blue) interact directly with this impurity site through a \(C_4\)-symmetric interaction.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.2\textwidth]{pWaveEsiam.pdf}
	\caption{General structure of an embedded extended SIAM, where the impurity site is a part of the 2D square lattice. The four nearest-neghbour sites (dark blue) interact with the impurity site through one or more processes, and can host local interactions. The other lattice sites (light blue) are completely non-interacting.}
	\label{embeddedEsiam}
\end{figure}
\subsection{Hamiltonian}
We consider an impurity site \(\vec S_d\) interacting with a two-dimensional tight-binding conduction bath through a highly localised interaction, described by the Hamiltonian
\begin{equation}\begin{aligned}
	\mathcal{H} = H_\text{cbath} + H_\text{imp} + H_\text{imp-cbath} + H_\text{cbath-int}~,
\end{aligned}\end{equation}
where \(H_\text{cbath}\) is the kinetic energy arising out of nearest-neighbour hopping processes and $H_\text{imp}$ is the particle-hole symmetric impurity-local term,
\begin{equation}\begin{aligned}
	H_\text{cbath} = -2t\sum_{{\bf k}}\left[\cos(ak_x) + \cos(ak_y)\right] c^\dagger_{{\bf k},\sigma}c_{{\bf k},\sigma}~,\quad H_\text{imp} = -\frac{U}{2}\left(n_{d\uparrow} - n_{d\downarrow}\right)^2~.
\end{aligned}\end{equation}
For the impurity-bath interaction \(H_\text{imp-cbath}\) and the local interaction \(H_\text{cbath-int}\) in the bath, we take the simplest choices consistent with the 2D square lattice geometry:
\begin{equation}\begin{aligned}
	H_\text{imp-cbath} &=\frac{1}{2} J\sum_{\sigma_1,\sigma_2}\sum_{Z} {\bf S}_d\cdot c^\dagger_{Z\sigma_1}{\boldsymbol \tau}_{\sigma_1,\sigma_2} c_{Z\sigma_2} - V\sum_{\sigma,Z}\left(c^\dagger_{d\sigma}c_{Z\sigma} + \text{h.c.}\right)~,\\
	H_\text{cbath-int} &= -\frac{W}{2}\sum_{Z} \left(n_{Z\uparrow} - n_{Z\downarrow}\right)^2 ~,
\end{aligned}\end{equation}
where \(Z\) sums over all nearest-neighbour sites of the impurity site (navy blue circles in Fig.~\ref{embeddedEsiam}, henceforth referred to as bath zeroth sites). The impurity interacts with the bath zeroth sites in \(H_\text{imp-cbath}\) through four single-particle hybridisation terms of strength $V$ and spin-exchange terms of strength $J$. The other term \(H_\text{cbath-int}\) represents local Hubbard interactions on the four zeroth sites, again of uniform strength.

The 2D structure of the impurity model and its distinction from the infinite-dimensional counterpart becomes apparent once we transform the Hamiltonian to momentum space.The conduction bath operators transform as
\begin{equation}\begin{aligned}
	c^\dagger_{{\bf r}} = \sum_{\bf k}e^{-i{\bf k}\cdot{\bf r}}c^\dagger_{\bf k}~,
\end{aligned}\end{equation}
where the momentum states $\bf k$ are the plane wave solutions of the tight-binding model obtained in the non-interacting limit: $U=J=0,V=t$. Transforming all bath operators leads to the following form:
\begin{equation}\begin{aligned}
	H_\text{imp-cbath} &=\frac{1}{2} \sum_{\sigma_1,\sigma_2,{\bf k}, {\bf k}^\prime} J_{{\bf k}, {\bf k}^\prime}{\bf S}_d\cdot c^\dagger_{{\bf k}\sigma_1}{\boldsymbol \tau}_{\sigma_1,\sigma_2} c_{{\bf k}^\prime\sigma_2}~,\\
	H_\text{cbath-int} &= -\frac{1}{2}\sum_{{\bf k}, {\bf k}^\prime, {\bf q}, {\bf q}^\prime} W_{{\bf k}, {\bf k}^\prime,{\bf q}, {\bf q}^\prime} c^\dagger_{\bf k}c_{{\bf k}^\prime}c^\dagger_{\bf q}c_{{\bf q}^\prime}~,
\end{aligned}\end{equation}
where the momentum-dependent couplings are defined as
\begin{equation}\begin{aligned}
	J_{{\bf k}, {\bf k}^\prime} &= \frac{J}{2}\left[\cos\left({\bf k}_x - {\bf k}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y\right)\right] ~,\\
	W_{{\bf k}, {\bf k}^\prime,{\bf q}, {\bf q}^\prime} &= W\left[\cos\left({\bf k}_x - {\bf k}^\prime_x + {\bf q}_x - {\bf q}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y + {\bf q}_y - {\bf q}^\prime_y\right)\right] ~.
\end{aligned}\end{equation}
where the subscripts \(x\) and \(y\) represent components of the momenta vector. This form shows that the Kondo coupling is no longer independent of the exchange momentum, and encodes information of the lattice geometry in it.
